Public Opinions on Sexual Education in Public Schools

### 2024-03-15 paper/paper.qmd
Update paper and references

---
title: "Public Opinions on Sexual Education in Public Schools"
subtitle: "An Analysis of Political, Educational, and Generational Influences in the United States from 1972 to 2022"
author: 
  - Wanling Ma
thanks: "Code and data are available at: https://github.com/WanlingMa/Favor-or-oppose-sex-education-in-public-schools"
date: today
date-format: long
abstract: "This paper examines the public's views on sexual education in U.S. public schools, drawing on data from the General Social Survey (GSS). Our analysis reveals widespread support for sexual education, though support varies significantly by political ideology, education level, and age, with Democrats, the highly educated, and younger people being the most supportive. These variations underscore the deep correlation between individual traits and attitudes towards sexual education, reflecting broader societal debates. In conclusion, the study highlights the critical role of inclusive perspectives in developing sexual education policies, suggesting that such education could serve as a unifying tool for public health and societal welfare."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

This study looks into how people feel about sexual education in public schools across the U.S., using data from the General Social Survey (GSS). We find a wide range of support for sexual education, but there are significant differences based on political views, education, and age. In particular, Democrats and individuals with higher levels of education tend to support sexual education more compared to Republicans and those with less education. Also, younger people show more support for sexual education, suggesting changes in social norms and the influence of modern values on education preferences.

@@

Our estimand focuses on the relationship between political affiliation, educational attainment, and age, and their collective impact on public opinions toward sexual education in public schools. We aim to precisely measure how these factors influence support or opposition to sexual education, serving as the core of our study. This approach allows us to identify and understand the patterns that form public sentiment on this debated topic.

The paper is organized into sections covering [Data], [Results], and [Discussions]. The Data [Data] section details our use of the GSS dataset, including how we selected and cleaned the data, and the challenges we faced, such as missing information and the transition to online survey methods. The [Results] section outlines our findings, showing how opinions on sexual education correlate with factors like political affiliation, education level, and age. The [Discussions] synthesizes our findings in the context of existing research, interpreting the trends we've observed and suggesting paths for future research to explore the dynamics we've identified.

@@

For our study, we retrieved data from the General Social Survey (GSS), a comprehensive and widely respected source of data that captures Americans' attitudes, behaviors, and attributes since 1972. The dataset includes variables such as respondents' age, their highest educational degree, political identification, opinions on sexual education in public schools, and the year of data collection. This diverse set of variables allows us to analyze trends and factors influencing views on sexual education in public schools over time.

@@

The demographic dataset, described in @tbl-demographic, focuses on basic respondent information including age, highest educational degree, and political identification, complemented by the year of data collection. This dataset lays the groundwork for understanding the demographic backdrop against which opinions on sexual education are formed.

@@

When the General Social Survey (GSS) capturing the dataset described in the previous section [Source Data], each variable is quantified to accurately reflect respondents' perspectives and backgrounds. For instance, opinions on sexual education are gathered through direct questioning, such as, "Would you be for or against sex education in the public schools?", with responses coded into predefined categories for detailed analysis [@gssCodebook]. Educational attainment is measured by the highest level of education respondents have completed, categorized from 'Less than high school' to 'Graduate degree'. Age is recorded as a continuous variable, offering insights into generational differences in attitudes towards sexual education. This systematic approach ensures that the dataset reliably mirrors respondents' characteristics of the American populations, providing a solid base for exploring the dynamics influencing public opinions on sexual education.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-ages
#| fig-cap: "Distribution of Respondent Ages"

@@

The 2022 General Social Survey (GSS) undertook a methodological transition to a multimode survey design, including face-to-face, web, and telephone interviews. This shift is primarily caused by the challenges presented by the COVID-19 pandemic [@gssCodebook]. This adaptation introduced an experimental design where sampled cases were randomly assigned to either start with face-to-face interviews followed by a web survey for non-respondents, or begin with a web survey and then offer face-to-face interviews for non-respondents. This experiment aimed to explore the impacts of survey mode on response patterns and data quality, with 94.2% of respondents completing the survey in a single mode and 5.8% in multiple modes. Notably, the survey achieved a 50.5% response rate, a significant increase from the 17% response rate of the primarily web-based 2021 GSS, underscoring the potential of multimode approaches to enhance participation [@gssCodebook].

The implications of this methodological shift for research, particularly studies like ours on public opinions regarding sexual education in public schools, are multifaceted. The introduction of different modes may introduce mode effects, where the method of data collection influences how participants respond to questions, especially on sensitive topics. This variability could lead to differences in response rates, the completeness of data, and possibly the willingness of respondents to discuss controversial issues like sexual education. For our survey, this methodological evolution represents a data limitation in that comparisons across years or interpretations of trends in public opinion must be approached with caution. Understanding that responses from 2022 might be influenced by the mode of data collection is crucial for accurately analyzing shifts in public opinion on sexual education, requiring sophisticated statistical techniques to disentangle genuine changes in attitudes from methodological artifacts.

@@

The General Social Survey (GSS) employs a split-ballot design, dividing questions among different versions of the questionnaire, or "ballots," to efficiently cover a wide array of topics. The absence of Ballot C from the Sexual Education Dataset, while Ballots A and B are present, indicates a notable gap in the dataset. This omission is significant because each ballot typically targets a distinct set of questions or themes, meaning that unique insights and data collected through Ballot C are missing from the analysis. The lack of Ballot C's data may lead to a loss of insights from respondents who were allocated to this ballot, potentially skewing the understanding of public opinions or demographic differences related to sexual education. This gap not only limits the breadth of questions that can be analyzed but also affects the study's ability to draw comprehensive conclusions about attitudes toward sexual education. By missing out on the data from Ballot C, researchers are deprived of a full picture, which can undermine the representativeness of findings and may lead to incomplete or biased interpretations of societal attitudes toward sexual education policies and perceptions.

@@

The absence of data for certain years, notably the shift from a planned 2020 data collection to 2021, presents a discontinuity in the longitudinal dataset of the GSS. This gap, particularly influenced by the COVID-19 pandemic, poses challenges for trend analysis and longitudinal studies aiming to understand changes in attitudes toward sexual education over time. The year 2020 was pivotal globally due to the pandemic, and missing data from this period means missing insights into how societal attitudes may have shifted in response to the global crisis and its impact on education systems, including sexual education. Transitioning to 2021 for data collection attempts to bridge this gap, but the temporal discontinuity introduces complexity in interpreting trends, requiring researchers to account for this break in the data when analyzing changes over time or drawing comparisons across years.

@@

The dataset's significant proportion of missing cases, with 31,615 cases not providing responses to questions related to sexual education out of a total of 72,390 cases, underscores a substantial data availability issue. This high rate of non-response could indicate several limitations, including possible respondent discomfort with the topic, lack of knowledge, or even survey fatigue. In the context of sexual education—a subject that can be sensitive for many individuals—the willingness or ability to answer related questions might not be evenly distributed across the survey population. This unevenness risks skewing the data towards those more inclined to respond, potentially those with stronger opinions or more direct experience with the topic, thereby compromising the representativeness of the dataset. Consequently, this limitation necessitates careful consideration and possibly the application of statistical adjustments or weighting to mitigate biases and ensure that findings accurately reflect the broader population's views on sexual education.

@@

This study has navigated the complex terrain of public opinion on sexual education in public schools, employing a robust dataset from the General Social Survey (GSS) that spans several decades. Despite the challenges posed by missing data in crucial years like 2020 and the shift to online survey methodology, the analysis has illuminated the influential role of political affiliations, educational levels, and generational perspectives on sexual education attitudes. The results underscore a notable trend: higher educational attainment and Democratic political leanings correlate with support for sexual education, while generational gaps reveal more conservative stances among older age groups.

The limitations of the study, including incomplete datasets and the lack of qualitative depth in multiple-choice responses, suggest avenues for richer, more nuanced future research. By integrating open-ended survey questions and employing the capabilities of Large Language Models for text analysis, future work can capture the complexities behind the statistics. The pursuit of a broader, more inclusive range of data promises to enrich the analysis, providing the multidimensional insight required to inform effective policy and educational program development.

In the end, our findings reflect not just the collective pulse on sexual education in schools but also the broader ideological and cultural currents shaping American society. This paper lays the groundwork for ongoing discussions and research that will continue to refine our understanding of public opinion and contribute to the creation of educational environments that honor the diverse needs and values of students, parents, and educators alike.

>>>>

---
title: "Public Opinions on Sexual Education in Public Schools"
subtitle: "An Analysis of Political, Educational, and Generational Influences in the United States from 1972 to 2022"
author: 
  - Wanling Ma
thanks: "Code and data are available at: https://github.com/WanlingMa/Favor-or-oppose-sex-education-in-public-schools"
date: today
date-format: long
abstract: "This paper examines the public's views on sexual education in U.S. public schools, drawing on data from the General Social Survey (GSS). Our analysis reveals that the support for sexual education varies significantly by political ideology, education level, and age, with Democrats, the highly educated, and younger people being the most supportive. These variations underscore the deep correlation between individual traits and attitudes towards sexual education, reflecting broader societal debates. In conclusion, the study highlights the critical role of inclusive perspectives in developing sexual education policies, suggesting that such education could serve as a unifying tool for public health and societal welfare."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

This paper looks into how people feel about sexual education in public schools across the U.S., using data from the General Social Survey (GSS) from 1972 to 2022. We find a wide range of support for sexual education, but there are significant differences based on political views, education, and age. In particular, Democrats and individuals with higher levels of education tend to support sexual education more compared to Republicans and those with less education. Also, younger people show more support for sexual education, suggesting changes in social norms and the influence of modern values on education preferences.

@@

Our estimand is the relationship between political affiliation, educational attainment and age, on public opinions toward sexual education in public schools. We aim to measure how these factors influence support or opposition to sexual education, serving as the core of our study. This approach allows us to identify and understand the patterns that form public sentiment on this debated topic.

The paper is organized into sections covering [Data], [Results], and [Discussions]. The [Data] section details our use of the GSS dataset, including how we selected and cleaned the data, and the challenges we faced, such as missing information and the transition to online survey methods. In the [Results] section, we present our key findings, illustrating the correlations between opinions on sexual education and various factors including political affiliation, education level, and age. Finally, the [Discussions] section integrates our findings with existing literature, offering interpretations of observed trends and proposing directions for future research that build on the identified dynamics.

@@

For our study, we retrieved data from the General Social Survey (GSS), a comprehensive and widely respected source of data that captures Americans' attitudes, behaviors, and attributes since 1972 [@GSS2022]. The dataset includes variables such as respondents' age, their highest educational degree, political identification, opinions on sexual education in public schools, and the year of data collection. This diverse set of variables allows us to analyze trends and factors influencing views on sexual education in public schools over time.

@@

The demographic dataset, described in @tbl-demographic, focuses on basic information about respondents, including age, highest educational degree, and political identification, complemented by the year of data collection. This dataset lays the groundwork for understanding the demographic backdrop against which opinions on sexual education are formed.

@@

When the General Social Survey (GSS) capturing the dataset described in the previous section, each variable is quantified to accurately reflect respondents' perspectives and backgrounds. For instance, opinions on sexual education are gathered through direct questioning, such as, "Would you be for or against sex education in the public schools?", with responses coded into predefined categories for detailed analysis [@gssCodebook]. Educational attainment is measured by the highest level of education respondents have completed, categorized from 'Less than high school' to 'Graduate degree'. Age is recorded as a continuous variable, offering insights into generational differences in attitudes towards sexual education. This systematic approach ensures that the dataset reliably mirrors respondents' characteristics of the American populations, providing a solid base for exploring the dynamics influencing public opinions on sexual education.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-ages
#| fig-cap: "Distribution of Respondent Ages in the GSS Dataset Over 1974 - 2022"

@@

In 2022, the General Social Survey (GSS) transitioned to a multimode survey design combining face-to-face, web, and telephone interviews in response to COVID-19. This methodological shift employed an experimental design, assigning respondents randomly to different initial survey modes to examine the impacts on response patterns and data quality. Notably, 94.2% of participants completed the survey in a single mode, and the overall response rate surged to 50.5% from 2021's 17%, demonstrating the potential of multimode strategies to enhance survey participation.

This shift introduces specific challenges for analyzing public opinions on sensitive topics like sexual education. Mode effects, resulting from how different survey mediums impact participant responses, can obscure the true nature of public sentiment, undermining the reliability of findings. For instance, respondents might be inclined to provide untruthful answers in web surveys or opt for neutral responses like "Don't know" in face-to-face settings. These variations necessitate careful consideration in data analysis, particularly in ensuring that observed changes in opinions are genuine reflections of public sentiment, not merely consequences of the survey's methodological changes.

@@

The General Social Survey (GSS) uses a split-ballot design to cover various topics, but the absence of Ballot C in the Sexual Education Dataset creates a significant gap. This omission removes potentially unique insights since each ballot targets specific questions or themes. The missing Ballot C data could skew interpretations of public opinion or demographic differences regarding sexual education, limiting the analysis scope and hindering our ability to draw well-rounded conclusions. Consequently, the lack of Ballot C compromises the dataset's comprehensiveness, potentially leading to biased or incomplete understandings of public attitudes towards sexual education policies.

@@

The shift in GSS data collection from 2020 to 2021, influenced by the COVID-19 pandemic, introduces a gap in the longitudinal dataset, complicating trend analysis and the study of evolving attitudes towards sexual education. Missing data from the pivotal year of 2020 limits insights into societal attitude shifts during the pandemic and its effects on education, including sexual education. While data collection in 2021 aims to fill this void, the discontinuity complicates trend interpretation, necessitating careful consideration of this break by researchers analyzing longitudinal changes or making cross-year comparisons.

@@

The dataset reveals a notable data availability issue, with 31,615 out of 72,390 cases lacking responses on sexual education, hinting at challenges like respondent discomfort, lack of knowledge, or survey fatigue. Given the sensitivity of sexual education, response willingness varies, potentially biasing results towards those more likely to participate, such as individuals with stronger opinions or direct experiences. This discrepancy threatens the dataset's representativeness, requiring meticulous analysis and potential statistical adjustments to align findings with the general population's views accurately.

@@

This study offers a concise exploration of public attitudes towards sexual education in public schools, leveraging the extensive General Social Survey dataset. Despite data gaps, notably from 2020, and a shift to online methodologies, our findings clearly demonstrate that higher education attainment and Democratic leanings enhance support, while older generations tend to be more conservative.

Acknowledging the limitations of our approach, including the dataset's incompleteness and the challenges posed by multiple-choice data, we see a pathway for deeper inquiry. Future research could benefit from incorporating open-ended questions and advanced text analysis to uncover the nuanced reasons behind public opinions. This study's strength lies in its use of a comprehensive, longitudinal dataset and its innovative analysis, setting a solid foundation for future work to build upon.

Ultimately, our research not only maps the current landscape of opinions on sexual education but also highlights the evolving societal norms affecting this crucial aspect of education. By pointing towards the need for policies and programs that reflect the complexity of public sentiment, this study contributes valuable insights into the ongoing dialogue around sexual education, advocating for an educational approach that resonates with a diverse society.
